---
title: "Bioinformatics in R: Statistical Learning very brief intro"
author: "J. Cesar Ignacio Espinoza - Cesar   "
date: "Week 07: April 29th and May 1st 2024"
output: 
  html_document: 
    highlight: espresso
    theme: cerulean
---

## Statistical Learning is Machine Learning

After the lecture in class we will be running some simple linear models.

```{r}
#### First install the car package, this has a bunch of classic datasets used when learning ML.
#install.packages("car")
library(car)
library(ggplot2)
library(readr)
```

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.

```{r}
#### Look at the very elitist prestige package
help("Prestige")
```

### First Model 

```{r}
b <- coef(reg1)[1]
m <- coef(reg1)[2]

ggplot(Prestige, aes(x= education, y = prestige)) + geom_point() + geom_abline(intercept = a, slope = b, color = "blue")
```

# prestige = b + m * education.

```{r}
reg1<-lm(prestige ~ education, data = Prestige)
summary(reg1)
```
HINT: You can get the weights of the model (coefficients) by doing this:

intercept <- coef(reg1)[1]
slope <- coef(reg1)[2]
You can add them to your ggplot with another geometry.

```{r}
### Look at the example below, we are adding multiple predictors
reg2<-lm(prestige~education+log(income)+women, data=Prestige)
summary(reg2)
```

```{r}
### We can add categorical data.
reg3<-lm(prestige~education+log(income)+type,data=Prestige)
summary(reg3)
```

```{r}
#reorder categorical data to have a "basal" category 
Prestige$type <- with(Prestige, factor(type, levels = c("bc","wc","prof")))
```

# Now let's try with Genetic data, upload the dataset from canvas.


```{r setup}
    ### Edit this line so your notebook renders properly
    knitr::opts_knit$set(root.dir = normalizePath("C:/Users/Juchy/OneDrive/Documents/School/School Work/Bioinformatics in R")) 
```

```{r}
data <- read_csv("Cholesterol.csv")
```

Simple excercise, which snps are significantly associated with Cholesterol level in blood?

```{r}
### Plot the data, 

ggplot(data, aes(x= rs7527051, y = CHOL)) + geom_point()

snp1<-lm(rs6669795 ~ CHOL, data = data)
snp2<-lm(rs7527051 ~ CHOL, data = data)
snp3<-lm(rs12140539 ~ CHOL, data = data)
snp4<-lm(rs2250402 ~ CHOL, data = data)
snp5<-lm(rs9509213 ~ CHOL, data = data)
snp6<-lm(rs12142199 ~ CHOL, data = data)
snp7<-lm(rs9782908 ~ CHOL, data = data)

summary(snp1)

# Adjusted R squared value of 0.011 means that 1.1% of cholesterol is explained by this SNP.

```


